### 2022-08-24 article.tex
Progress

From this simple example, it will be easier
to measure the level of IL6RA in the blood, than using a gene panel.
However, there are associations published for many diseases,
in which multiple SNPs contribute to being more likely 
to develope a disease in the future. Here, the phenotype (having 
a disease in the future) is impossible to detect at the present
and associations found in earlier studies are used to create a gene panel.
As creating a gene panel is costly, those associations better be correct.

@@

The experiment described above is run by code. 
It was code that detected the relationship between the genotype
(in this case, SNP \verb|rs12133641|) 
and the phenotype (in this case,the concentration of IL6RA).
To detect this association, 
there was no fieldwork, nor lab-work involved.
For this experiment, 
it has been the researcher that has written the code,
after which that code does the actual work. 
A genetic epidemiologist does not need a lab, but computational power instead.

To \emph{obtain} the raw genetic data, yes, there may be fieldwork and/or lab-work
involved. For example, a researcher needs to go into the field to take
samples (e.g. blood) of humans or other species. To analyse these samples,
a researcher needs a lab. Depending on the technique being used on
how to analyse the samples, there may be a big bioinformatics step
to aggregate the measurements into useful genetic data. 
Although this paper does not focus on the code run to 
aggragate the raw data into useful genetic data, the same arguments
can be made for that code as in the example: that code is 
the ground truth of how the raw data is collected into useful genetic data.

>>>>

From this simple example, it will be easier
to measure the level of IL6RA in the blood, than using a gene panel,
as blood tests are easier and cheaper.
However, there are associations published for many diseases,
in which one (e.g. phenylketonuria) or many (e.g. \cite{bruce2009metabolic}) 
SNPs contribute to being more likely to develope a disease in the future. 
Here, the phenotype (having a disease in the future) 
is impossible to detect at the present
and associations found in earlier studies are used to create a gene panel.
As creating a gene panel is costly, those associations better be correct.

@@

The experiment described above is run by code. 
It was code that detected the relationship between the genotype
(in this case, SNP \verb|rs12133641|) 
and the phenotype (in this case, the concentration of IL6RA).
To detect this association, 
there was no fieldwork nor lab-work involved.
For this experiment, 
it has been the researcher that has written the code,
after which that code does the actual work. 
A genetic epidemiologist does not need a lab, but computational power instead.

To \emph{obtain} the raw genetic data, there may have been fieldwork and/or lab-work
involved. For example, a researcher needs to go into the field to take
samples (e.g. blood) of humans or other species. 
To analyse these samples,
a researcher needs a lab. 
Depending on the analysis, there may be a big bioinformatics step
to aggregate the measurements into usable data. 
Although this paper does not focus on the code run to 
aggragate the raw data into useful genetic data, the same arguments
can be made for that code as in the example: that code is 
the ground truth of how the raw data is collected into useful genetic data.
